---
title: "Distribution_TT"
output: html_document
date: "2025-03-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# A Distribution-Free test for the Slope $\beta$

Data: $(x_i,y_i),i=1,2,...,n$. The $X_i's$ are assumed to be distinct, without loss of generality, we take $x_1<x_2<...<x_n$

Assumptions:

1. $y_i = \alpha+\beta x_i + \epsilon_i, i=1,2,...,n$
2. $\epsilon_i \sim \text{ iid } F(unknown,continuous)$, $med(F)=0$, $F^{-1}(1/2)=0$

# Test for $\beta$:

$$H_0:\beta = \beta_0, H_1: \beta > \beta_0$$

Let $d_i=y_i-\beta_i x_i,i=\sqrt{n}$.

The Theil Statistic is defined by:

$$T_t = \sum_{i<j} sgn(d_j-d_i)=\sum^{n-1}_{i=1}\sum^{n}_{j=i+1}sgn(d_j-d_i)$$

Let us find the distribution of $T_t$

Steps:

1. Generate random samples of $(X,Y)$
2. Calculate $d_i$
3. Calculate $T_s$
4. Repeat steps 1-3 10000 times.
```{r}
# ---------- Random Samples of (X,Y) ----------
sampleMax <- 1000
sampleSize <- 4
xSamples <- c(1:sampleMax)
alpha <- 1 #runif(n = 1, min = 1, max = 10) # Value of Alpha
beta <- 1 #runif(n = 1, min = 1, max = 10) # Value of Beta

# ---------- Residuals ----------
data <- sample(xSamples, sampleSize, replace = TRUE)
resid <- data - median(data)
#cat("Median of Residuals: ", median(resid))

# ---------- Make Data and Line ----------
x_i <- sample(xSamples, sampleSize, replace = TRUE) # Random Sample of X_i (sorted)
y_i <- alpha + beta * x_i + resid # Y_i

# ---------- Plot (X,Y) ----------
plot(x_i, y_i)

# ---------- Calculate d_i and signs ----------
d_i <- y_i - beta * x_i # Calculate d_i

# ---------- Calculate T_t ----------
sumVec <- c()
for(i in 1:(sampleSize-1)){
  for(j in (i+1):(sampleSize)){
    #cat(j,",",j+1,",",i,",",sampleSize - i)
    sign <- sign(d_i[j] - d_i[i])
    sumVec <- append(sumVec, sign)
    #print(sign)
  }
}

d_i
sum(sumVec)
```

Let's find the Null Distribution of $T_t$ for a specific sample size. We will use random values of $\alpha$ and $\beta$, create residuals with median $0$, create the data $(X_i,Y_i)$ where $i = 1,2,..n$, and create Kendall Statistics for 10000 iterations.
```{r}
iterations <- 10000
superSumVec <- c()
sampleMax <- 1000
sampleSize <- 4
xSamples <- c(1:sampleMax)

for(k in 1:iterations){
  #---------- Alpha and Beta Values ----------
  alpha <- runif(n = 1, min = 1, max = 100) # Value of Alpha
  beta <- runif(n = 1, min = 1, max = 100) # Value of Beta
  
  # ---------- Residuals ----------
  data <- sample(c(1:sampleMax), sampleSize, replace = TRUE)
  resid <- data - median(data) # Create Residuals with MEdian = 0
  #cat("Median of Residuals: ", median(resid))

  # ---------- Make Data ----------
  x_i <- sample(xSamples, sampleSize, replace = TRUE) # Random Sample of X_i
  y_i <- alpha + beta * x_i + resid # Y_i

  # ---------- Calculate d_i and signs ----------
  d_i <- y_i - beta * x_i # Calculate d_i
  signs <- sign(d_i) # Convert to Signs

  sumVec <- c()
  for(i in 1:(sampleSize-1)){
    for(j in (i+1):(sampleSize)){
      #cat(j,",",j+1,",",i,",",sampleSize - i)
      sign <- sign(d_i[j] - d_i[i])
      sumVec <- append(sumVec, sign)
      #print(sign)
    }
  sum <- sum(sumVec)   
  }
  superSumVec <- append(superSumVec, sum)
}

hist(superSumVec, breaks = 50, col = "lightblue", main = "Histogram of Kendall's Ks Statistic")
```


Let's make a column of the table
```{r}
x <- seq(0, 100, 2) # Sequence of x's from the Table
probVec <- numeric(length(x)) # Empty Vector of Probabilities

for(i in 1:length(x)){
  prob <- length(superSumVec[superSumVec >= x[i]])/iterations
  probVec[i] <- prob 
}

round(probVec, 3)

```

# Creating the Table using a Function

Let's make the Kendall Null Distribution a function
```{r}
kendallNull <- function(sampleSize){
  # ---------------------------------------------
  # ---------- Null Distribution of Ks ----------
  # ---------------------------------------------
  iterations <- 10000
  superSumVec <- c()
  sampleMax <- 1000
  sampleSize <- sampleSize
  xSamples <- c(1:sampleMax)

  for(k in 1:iterations){
    #---------- Alpha and Beta Values ----------
    alpha <- runif(n = 1, min = 0, max = 1) # Value of Alpha
    beta <- runif(n = 1, min = 0, max = 1) # Value of Beta
  
    # ---------- Residuals ----------
    data <- sample(c(1:sampleMax), sampleSize, replace = TRUE)
    resid <- data - median(data) # Create Residuals with MEdian = 0
    #cat("Median of Residuals: ", median(resid))

    # ---------- Make Data ----------
    x_i <- c(1:sampleSize) # x_i
    #x_iSample <- sample(xSamples, sampleSize, replace = TRUE) # Random Sample of   X_i
    y_i <- alpha + beta * x_i + resid # Y_i

    # ---------- Calculate d_i and signs ----------
    d_i <- y_i - beta * x_i # Calculate d_i
    signs <- sign(d_i) # Convert to Signs
  
    # ---------- Calculate Ks Statistic ----------
    sumVec <- c()
    for(i in 1:(sampleSize-1)){
      for(j in (i+1):(sampleSize)){
        #cat(j,",",j+1,",",i,",",sampleSize - i)
        sign <- sign(d_i[j] - d_i[i])
        sumVec <- append(sumVec, sign)
        #print(sign)
      }
    sum <- sum(sumVec)   
    }
    superSumVec <- append(superSumVec, sum)
  }
  x <- seq(0, 100, 2) # Sequence of x
  probVec <- c() # Empty Vector of p-values

  # ---------------------------------------------------
  # ---------- Calculate Probability Ks >= x ----------
  # ---------------------------------------------------
  for(i in 1:length(x)){
    prob <- length(superSumVec[superSumVec >= x[i]])/iterations # p-value
    probVec <- append(probVec, prob) # Append p-value Vector
  }
  
  # ---------------------------------------------------
  # ---------- Create Table of Distributions ----------
  # ---------------------------------------------------
  table <- round(probVec, 3)
  return(table)
}

kendallNull(4)
```

# Making Table A.30

Let's make the entire table into a dataframe
```{r}
n <- c(4,5,8,9,12,13,16,17,20,21,24,25,28,29,32,33,36,37,40) # Sample Sizes
x <- seq(0, 100, 2) # x where P(K >= x)

startTime <- Sys.time()
table <- sapply(n, kendallNull) # Make the Table of P-Values
df <- data.frame(x, table) # Make a Dataframe for the P-Values
colnames(df) <- c("x", n)
endTime <- Sys.time()

endTime - startTime

df
```

Export the Table to a CSV File
```{r}
write.csv(df, "C:/Users/EJ's/Desktop/Kendall_pValues2.csv" )
print("CSV File Created Successfully")
```


