---
title: "Pitman Test Null Distribution"
author: "Edgar Derricho"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Run All First

- You will need to imput your dataset and the proposed median in lines 45 and 46.

# Pitman Test

Suppose that $X_1,X_2,...,x_n \sim iid \text{ }F$ where $F$ is symmetric aboit its mean (median) $\theta$. We want to test the hypothesis that $H_0:\theta = \theta_0(\text{ known value}),H_1 \theta \neq \theta_0$.

Method:

consider the $n$ differences: $d_i = x_i-\theta_0,i=1,2,...,n$. Under $H_0$, the sign of each $d_i$ is equally likely to be positive or negtive.

ie. $P(d_i>0)=P(d_i<0)=\frac{1}{2}$

Let

$$S_+=\sum^{n}_{i=1}d_iI(d_i>0),S_-=+\sum^{}_{i=1}|d_i|I(d_i<0)$$

**Note: $S_+$ is the sum of the positive deviations from $\theta_0$**

Larger observed values of $S_+$ indicate that $H_0$ is not true (clearly when $H_0$ does not hold there is more likely to be a preponderance wither of positive or negative signs assiciated with the $d_i$)

Reject $H_0$ at level $\alpha$ when

p-value = $P(S_+ \geq S_{+obs}) < \alpha$ where $S_{+obs}$ is an observed value of $S_+$

# Null Distribution of Pitman Statistic $S_+$

We will simulate 10000 pitman tests statistics to find the null distribution of $S_+$

You will need to input your sample and $\theta_0$
```{r}
# Sample
sample <- c(12,18,24,26,37,40,42,47,49,49,78,108) # Sample
theta <- 30 # Theta_0
```

Pitman Null Distribution Calculation
```{r}
# ---------- Sample Sizes and Possible Values ----------
n <- length(sample) # Length of Sample
diff <- abs(sample - theta) # Difference Vector
sPlusVector <- numeric(iterations) # Empty Vector of SPlus Values
sMinusVector <- numeric(iterations) # Empty Vector of SMinus Values
indexes <- c(1:n) 

# ---------- Pitman Tests ----------
iterations <- 10000
for(i in 1:iterations){
  ones <- rep(1, n) # Vector of 1's to Multiply
  signs <- sample(x = c(1,-1),
                  size = n,
                  replace = TRUE) # Number of Negative Signs
  modDiff <- diff * signs # Sample with Negative Indexes
  sum <- sum(modDiff[modDiff > 0]) # Sum of Modified Difference Vector
  sumNeg <- abs(sum(modDiff[modDiff < 0])) # Sum of S Minus Vector
  sPlusVector[i] <- sum # Append SPlus Vector with SPlus
  sMinusVector[i] <- sumNeg # Append SMinus Vector with SMinus
}

#---------------------------------------------------------------
# ---------- Right-Tail Critical Values for SPlus ----------
ninetyPercent <- quantile(sPlusVector, 0.90)
ninetyFivePercent <- quantile(sPlusVector, 0.95)
ninetyNinePercent <- quantile(sPlusVector, 0.99)

# ---------- Right-Tail Critical Values for SPlus ----------
ninetyPercentN <- quantile(sMinusVector, 0.90)
ninetyFivePercentN <- quantile(sMinusVector, 0.95)
ninetyNinePercentN <- quantile(sMinusVector, 0.99)
# ---------------------------------------------------------------

# ----------------------------------------------------------------
# ---------- Print Critical Values ----------
cat("---------- For S-Plus ----------","\n",
  "Critical Value at 90%: ", ninetyPercent,"\n",
    "Critical Value at 95%: ", ninetyFivePercent,"\n",
    "Critical Value at 99%: ", ninetyNinePercent, "\n")

# ---------- Print Critical Values ----------
cat("---------- For S-Minus ----------","\n",
  "Critical Value at 90%: ", ninetyPercentN,"\n",
    "Critical Value at 95%: ", ninetyFivePercentN,"\n",
    "Critical Value at 99%: ", ninetyNinePercentN)
# -----------------------------------------------------------------

# ---------- Plot Histograms ----------
hist(sPlusVector)
hist(sMinusVector)
```

Notice the $S_+$ and $S_-$ distributions are virtually identical.

P-Value Calculator for Calculated $S_+$
```{r}
splus <- 95 # S-Plus from Problem
length(sMinusVector[sMinusVector>=splus])/iterations # P-Value
```

P-Value Calculator for Calculated $S_-$
```{r}
sminus <- 40 # S-Plus from Problem
length(sMinusVector[sMinusVector<=sminus])/iterations # P-Value
```

